---
title: 'STAT 206 Lab 2'
author: Vanessa Feng
date: 10/13/2019
output: pdf_document
---

**Due Monday, October 14, 5:00 PM**

***General instructions for labs***: Labs must be completed as a pdf file.  Give the commands
to answer each question in its own code block, which will also produce plots that will be
automatically embedded in the output file. Each answer must be supported by written statements
as well as any code used. 

***Agenda***: Manipulating data frames; practicing iteration; practicing re-writing code;
checking how reliable random methods are.

## Part I -- Data Frames

==================================

R includes a number of pre-specified data objects as part of its default installation. We will
load and manipulate one of these, a data frame of 93 cars with model year 1993. Begin by ensuring that you can load this data with the commands

```{r}
library(MASS)
data(Cars93)
```

#### Begin by examining the data frame with the command `View(Cars93)` to understand the
#### underlying object. You will need to use functions and other commands to extract elements
### for this assignment.
```{r, eval=FALSE}
View(Cars93)
```
### 1. Obtain a `summary()` of the full data structure. Can you tell from this how many rows
### are in the data? If so, say how; if not, use another method to obtain the number of rows.
```{r}
summary(Cars93)
# rows = 93
# Summary results get some information of each column, the column with 
# digital shows the digital features, and the one without digital shows
# the lable and the number of the lable. Thus, the number of rows equals
# to the sum of these lable's number.
# Such as column Manufacturer, 8+8+6+5+5+4+57=93, which is the number of the rows.
```
### 2. What is the mean price of a car with a rear-wheel drive train?
```{r}
Cars_rear = c()
for(i in 1:93){
  if(Cars93$DriveTrain[i] == "Rear"){
    Cars_rear <- c(Cars_rear, Cars93$Price[i])
  }
}
if(!is.null(Cars_rear)){
  cat("The mean price of a car with a rear-wheel drive train is :", mean(Cars_rear), "\n")
}else{
  cat("There is no information about the car with a rear-wheel drive train.")
}
```
### 3. What is the minimum horsepower of all cars with capacity for 7 passengers? With a capacity of at least 6 passengers?
```{r}
Cars_horsepower = c()
for(i in 1:93){
  if(is.integer(Cars93$Passengers) & Cars93$Passengers[i] == 7){
    Cars_horsepower <- c(Cars_horsepower, Cars93$Horsepower[i])
  }
}
if(!is.null(Cars_horsepower)){
  cat("The minimum horsepower of all cars with capacity for 7 passengers
      is :", min(Cars_horsepower), "\n")
}else{
  cat("There is no information about the car with capacity for 7 passengers.\n")
}

Cars_horsepower = c()
for(i in 1:93){
  if(is.integer(Cars93$Passengers) & Cars93$Passengers[i] >= 6){
    Cars_horsepower <- c(Cars_horsepower, Cars93$Horsepower[i])
  }
}
if(!is.null(Cars_horsepower)){
  cat("The minimum horsepower of all cars with capacity for at lease
      6 passengers is :", min(Cars_horsepower), "\n")
}else{
  cat("There is no information about the car with capacity at lease
      6 passengers.\n")
}
```
### 4. Assuming that these cars are exactly as fuel efficient as this table indicates, find 
### the cars that have the maximum, minimum and median distance travellable for highway
### driving.  You will need at least two columns to work this out; why those two?
```{r}
Cars_distance <- Cars93$MPG.highway * Cars93$Fuel.tank.capacity
max(Cars_distance)
min(Cars_distance)
median(Cars_distance)
# The fuel efficient of cars are shown in the column MPG.city and 
# MPG.highway of the table.
# MPG(miles per gallon) means the number of miles the car can run 
# using one gallon fuel.
# The distance travellable for highway driving = MPG.highway * cars' 
# fuel tank capacity.
```

## Part II -- Reproducibility and Functions

================================================

### Some of the lectures have included examples of planning production for a factory that 
### turns steel and labor into cars and trucks.  Below is a piece of code that optimizes the
### factory's output (roughly) given the available resources, using a `repeat` loop.  It's
### embedded in a function to make it easier for you to run.

```{r}
factory.function <- function (cars.output=1, trucks.output=1) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)
  
  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, randomly
     # to weak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
  }

  return(output)
}
```

### 5. Run the function above with the command

```{r}
result <- factory.function()
result
cat("The final result shows that the factory can get about", result[1], " cars
    and ", result[2], " trucks after maximizing the use of labors and steel.\n")
```

### to obtain a default output value, starting from a very low initial planned output. What
### is the final output capacity obtained?

### 6. Repeat this four more times to obtain new output values. Do these answers differ 
### from each other? If so why? If not, why not?
```{r}
factory.function()
factory.function()
factory.function()
factory.function()
# There are some difference among the answers.
# The difference is due to the the process of generating random numbers
# in the factory.function()
# When one of needed$labor and needed$steel exceeds its available amount,
# the output will increase or decrease a small random value 
# during(-0.1,0.1), to motivate the repeat process to continue.
# Thus, although there are some difference among the answers, 
# they are very small.
```
### 7. Right now, the number of `passes` is a value held within the function itself and not
### shared. Change the code so that the number of `passes` will be returned at the end of the
### function, as well as the final `output`. 
```{r}
factory.function <- function (cars.output=1, trucks.output=1) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)
  
  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       output <- list(output,  repeat_number=passes)
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, 
     # randomly to weak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
  }

  return(output)
}
factory.function()
```
### 8. Now, set the initial output levels to 30 cars and 20 trucks and run the code.  What 
### is the final output plan (`output`)?  What is the final demand for resources (`needed`)?
### Is the plan within budget and within the slack?  How many iterations did it take to
### converge (`passes`)? For all but `output` you will need to either print this message out
### deliberately, or return an object that contains all the quantities you want.
```{r}
factory.function <- function (cars.output=1, trucks.output=1) {
  factory <- matrix(c(40,1,60,3),nrow=2,
    dimnames=list(c("labor","steel"),c("cars","trucks")))
  available <- c(1600,70); names(available) <- rownames(factory)
  slack <- c(8,1); names(slack) <- rownames(factory)
  output <- c(cars.output, trucks.output); names(output) <- colnames(factory)
  
  passes <- 0 # How many times have we  been around the loop?
  repeat {
     passes <- passes + 1
     needed <- factory %*% output # What do we need for that output level?
     # If we're not using too much, and are within the slack, we're done
     if (all(needed <= available) &&
         all((available - needed) <= slack)) {
       output <- list(output=output,  repeat_number=passes, needed=needed)
       break()
     }
     # If we're using too much of everything, cut back by 10%
     if (all(needed > available)) {
       output <- output * 0.9
       next()
     }
     # If we're using too little of everything, increase by 10%
     if (all(needed < available)) {
       output <- output * 1.1
       next()
     }
     # If we're using too much of some resources but not others, 
     # randomly to weak the plan by up to 10%
      # runif == Random number, UNIFormly distributed, not "run if"
     output <- output * (1+runif(length(output),min=-0.1,max=0.1))
  }

  return(output)
}
```
```{r}
result <- factory.function(30, 20)
result
# According to the output
cat("We can finally produce ", result$output[1], " cars and ", 
    result$output[2], "trucks.\n")
cat("Producing these we need ", result$needed[1], " labors and 
    amount of", result$needed[2],
    "steel.\n")

available <- c(1600,70)
slack <- c(8,1)
if(all(result$needed <= available)){
  cat("Labor usage:", result$needed[1], " < 1600 and 
      steel usage: ", result$needed[2], "< 70\n")
  print("So, the plan within the budget.")
}else{
  print("The plan doesn't within the budget.")
}
if(all((available-result$needed) <= slack)){
  cat("Labor remain:", available[1]-result$needed[1], " < 8 and 
      steel remain: ", available[2]-result$needed[2], "< 1\n")
  print("The plan within the slack.")
}else{
  print("The plan doesn't within the slack.")
}
```
