---
title: 'STAT 206 Lab 7'
author: Xin Feng(Vanessa)
date: 11/16/2019
output: pdf_document
---

**Due Monday, November 18, 5:00 PM**

***General instructions for labs***: Labs must be completed as a pdf file.  Give the commands to answer each question in its own code block, which will also produce plots that will be automatically embedded in the output file. Each answer must be supported by written statements as well as any code used. 

***Agenda***: Debugging and testing

Testing for Outliers
==========

Identifying outliers in data is an important part of statistical analyses. One simple rule of thumb (due to John Tukey) for finding outliers is based on the quartiles of the data: the first quartile $Q_1$ is the value $\ge 1/4$ of the data, the second quartile $Q_2$ or the median is the value $\ge 1/2$ of the data, and the third quartile $Q_3$ is the value $\ge 3/4$ of the data. The interquartile range, $IQR$, is $Q_3 - Q_1$. Tukey’s rule says that the outliers are values more than 1.5 times the interquartile range from the quartiles -- either below $Q_1 - 1.5 IQR$, or above $Q_3 + 1.5 IQR$.  Consider the data values
```{r}
x <- c(2.2, 7.8, -4.4, 0.0, -1.2, 3.9, 4.9, 2.0, -5.7, -7.9, -4.9,  28.7,  4.9)
```
We will use these as part of writing a function to identify outliers according to Tukey’s rule. Our function will be called `tukey.outlier`, and will take in a data vector, and return a Boolean vector, `TRUE` for the outlier observations and `FALSE` elsewhere.

1. Calculate the first quartile, the third quartile, and the inter-quartile range of `x`. Some built-in R functions calculate these; you cannot use them, but you could use other functions, like `sort` and `quantile`.
```{r}
qt_x <- quantile(x)
qt_1 <- qt_x[[2]]
qt_3 <- qt_x[[4]]
iqr <- qt_3-qt_1
cat("The first quartile =", qt_1, '\n')
cat("The third quartile =", qt_3, '\n')
cat("The interquartile range of x =", iqr, '\n')
# IQR(x)
```
2. Write a function, `quartiles`, which takes a data vector and returns a vector of three components, the first quartile, the third quartile, and the inter-quartile range. Show that it gives the right answers on `x`. (You do not have to write a formal test for quartiles.)
```{r}
quartiles <- function(x){
  qt <- quantile(x)
  qt_1 <- qt[[2]]
  qt_3 <- qt[[4]]
  return(c(qt_1, qt_3, qt_3-qt_1))
}
quartiles(x)
# The answer of testing given x is right.
```
3. Which points in `x` are outliers, according to Tukey’s rule, if any?
```{r}
tukey.ot <- function(x){
  qt <- quartiles(x)
  rg_max <- qt[2] + 1.5*qt[3]
  rg_min <- qt[1] - 1.5*qt[3]
  index <- which(x>rg_max | x<rg_min)
  data <- x[index]
  if(!length(data)) print("There isn't outlier in the vector.")
  else return(cbind(index, data))
}
tukey.ot(x)
```
4. Write a function, `test.tukey.outlier`, which tests the function `tukey.outlier` against your answer in the previous question. This function should return `TRUE` if `tukey.outlier` works properly; otherwise, it can either return `FALSE`, or an error message, as you prefer. (You can do the next problem first, if you find that easier.)
```{r}
library(testthat)
test.tukey.outlier <- function(x){
  expect_type(x, "double")
  expect_type(tukey.outlier(x), "logical")
  
  result <- tukey.outlier(x)
  qt <- quartiles(x)
  rg_max <- qt[2] + 1.5*qt[3]
  rg_min <- qt[1] - 1.5*qt[3]
  judge = 0
  for(i in 1:length(x)){
    if(x[i] < rg_min | x[i] > rg_max){
      if(result[i]==FALSE){
        judge = 1
        cat("Error: function tukey.outlier() is wrong. The ", i, "-th data 
            in x is outlier.")
      }
    }else{
      if(result[i]==TRUE){
        judge = 1
        cat("Error: function tukey.outlier() is wrong. The ", i, "-th data 
            in x is not outlier.")
      }
    }
  }
  ifelse(!judge, return(TRUE), return(FALSE))
}
# Test: initial vector x
try(test.tukey.outlier(x))
# Test: if vector x contains character item.
try(test.tukey.outlier(c("a",1,4,5,7,9)))
```
5. Write `tukey.outlier`, using your `quartiles` function. The function should take a single data vector, and return a Boolean vector, take in a data vector, and return a Boolean vector, `TRUE` for the outlier observations and `FALSE` elsewhere. Show that it passes `test.tukey.outlier`.
```{r}
tukey.outlier <- function(x){
  qt <- quartiles(x)
  rg_max <- qt[2] + 1.5*qt[3]
  rg_min <- qt[1] - 1.5*qt[3]
  judge <- vector(mode="logical", length=length(x))
  judge[which(x>rg_max |x<rg_min)]=TRUE
  return(judge)
}
result <- tukey.outlier(x)
result
```
6. Which data values should be outliers in `-x`?
```{r}
tukey.ot(-x)
tukey.outlier(-x)
```
7. Which data values should be outliers in `100*x`?
```{r}
tukey.ot(100*x)
tukey.outlier(100*x)
```
8. Modify `test.tukey.outlier` to include tests for these cases.
```{r}
test.tukey.outlier <- function(x){
  expect_type(x, "double")
  expect_type(tukey.outlier(x), "logical")
  
  judge.x <- function(x){
    result <- tukey.outlier(x)
    qt <- quartiles(x)
    rg_max <- qt[2] + 1.5*qt[3]
    rg_min <- qt[1] - 1.5*qt[3]
    judge = 1
    for(i in 1:length(x)){
      if(x[i] < rg_min | x[i] > rg_max){
        if(result[i]==FALSE){
          judge = 0
          cat("Error: function tukey.outlier() is wrong. The ", i, "-th data 
              in",x," is outlier.")
        }
      }else{
        if(result[i]==TRUE){
          judge = 0
          cat("Error: function tukey.outlier() is wrong. The ", i, "-th data 
              in ",x," is not outlier.")
        }
      }
    }
    ifelse(judge, return(TRUE), return(FALSE))
  }
  judge <- judge.x(x) + judge.x(-x) + judge.x(100*x)
  ifelse(judge, return(TRUE), return(FALSE))
}
```
9. Show that your `tukey.outlier` function passes the new set of tests, or modify it until it does.
```{r}
tukey.outlier <- function(x){
  expect_type(x, "double")
  qt <- quartiles(x)
  rg_max <- qt[2] + 1.5*qt[3]
  rg_min <- qt[1] - 1.5*qt[3]
  judge <- vector(mode="logical", length=length(x))
  judge[which(x>rg_max |x<rg_min)]=TRUE
  return(judge)
}
tukey.outlier(x)
try(test.tukey.outlier(x))
# Test, input character
try(tukey.outlier(c('a',1,3,4,5,6)))
try(test.tukey.outlier(c('a',1,2,3,4)))
```
10. According to Tukey’s rule, which points in the next vector are outliers? What is the output of your function? If they differ, explain why.
```{r}
y <- c(11.0, 14.0, 3.5, 52.5, 21.5, 12.7, 16.7, 11.7, 10.8, -9.2, 12.3, 13.8, 11.1)
```

```{r}
tukey.ot(y)

tukey.outlier(y)
# The output of function tukey.outlier is a Boolean vector, `TRUE` for the 
# outlier observations and `FALSE` for the data which is not outlier.
# Due to the different output format, they seem difference, in fact, the 
# results are same.
```

